### 2022-11-04 initial_report.Rmd

---
title: "Initial report"
output: html_document
date: '`r Sys.Date()`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Read data
raw <- readr::read_tsv("data/ECAW_raw_data.tsv")
```

# Reproducibility and trustworthiness

The reproducible and trustworthiness ratings with the current system.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
typically_likert_data <-
  raw %>% 
    select(starts_with("typically_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
            "Strongly disagree",
            "Somewhat disagree",
            "Neither agree nor disagree",
            "Somewhat agree",
            "Strongly agree",
            "Missing",
            "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

typically_missing_data <-
  typically_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

typically_likert_data %>% 
  likert_plot_odd(.,
                  title = "Typically, studies that analyze preexisting\nobservational datasets, are ...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question"),
                  positive_side = "lower",
                  text_push = 0.05) +
  geom_text(typically_missing_data, mapping = aes(x = type, y = 1.15, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19)
    )
```

The reproducible and trustworthiness ratings of ECAW compared to the current system.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
ecaw_likert_data <-
  raw %>% 
    select(starts_with("ecaw_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more",
          "Missing",
          "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

ecaw_missing_data <-
  ecaw_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

ecaw_likert_data %>% 
  likert_plot_odd(.,
                  title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question"),
                  positive_side = "lower",
                  text_push = 0.05) +
  geom_text(typically_missing_data, mapping = aes(x = type, y = 1.15, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# ALSPAC

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 22, fig.height = 8}
alspac_likert_data <-
  raw %>% 
    select(starts_with("likert_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "type",
      values_to = "scale",
    ) %>%
    mutate(
      type = str_remove(type, "^[^_]*_")
    ) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree",
          "Missing",
          "I don't understand the question",
          "Unsure")
            ),
      type = case_when(
        type == "study" ~ "ALSPAC should\nrun a study",
        type == "prefer_ecaw" ~ "I would\nprefer ECAW",
        type == "opt_in" ~ "I would opt in",
        type == "less_willing" ~ "I would be\nless willing to use",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

alspac_missing_data <-
  alspac_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question", "Unsure")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing",
      scale == "Unsure" ~ "Unsure"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

alspac_likert_data %>% 
  likert_plot_odd(.,
                  title = "Thinking about a study you may run with ALSPAC data ...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question", "Unsure"),
                  positive_side = "lower",
                  text_push = 0.08,
                  limits = c(-1, 1)) +
  geom_text(alspac_missing_data, mapping = aes(x = type, y = 1.2, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# Current methods

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
method_data <-
  raw %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind analysis",
        type == "confirmatory" ~ "Confirmatory analysis",
        type == "exploratory" ~ "Exploratory analysis",
        type == "preregistered" ~ "Preregistration",
        type == "script" ~ "Analysis script",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

# method_missing_data <-
#   method_data %>% 
#   filter(scale %in% c("Missing", "I don't understand the question")) %>% 
#   mutate(
#     scale = case_when(
#       scale == "I don't understand the question" ~ "Don't understand",
#       scale == "Missing" ~ "Missing"),
#     text = paste0(scale, ": ", n)
#     ) %>% 
#   group_by(type) %>% 
#   summarise(text = stringr::str_c(text, collapse = "\n"))

method_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  labs(
    title = "The studies using preexisting observational data that I am involved in...",
    x = "Percentage",
    fill = "Response levels"
  ) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

# How many respondents are open to  be interviewed?

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8}
raw %>% 
  replace_na(list(contact = "Missing")) %>% 
  count(contact) %>% 
  mutate(
    contact = case_when(
      str_detect(contact, "Yes.") ~ "Yes",
      TRUE ~ contact),
    contact = factor(contact, levels = c("Yes", "No", "Missing")),
    sum_n = sum(n),
    percentage = n / sum_n,
    text_label = paste0(n, "/", sum_n)
  ) %>%
  ggplot() +
  aes(
    x = contact,
    y = percentage
  ) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = text_label), size = 7, vjust = -1) +
  scale_y_continuous(expand = c(0, 0), labels = scales::label_percent(), limits = c(0, 0.55)) +
  labs(
    x = "May the research team contact you about potential focus groups?",
    y = "Percentage"
  ) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 15)
  )
```

# Concerned with rigour and reproducibility on the field

```{r, echo = FALSE, message = FALSE, warning = FALSE, , fig.width = 20, fig.height = 8}
concerned_data <-
  raw %>% 
  replace_na(list(concerned = "missing")) %>%
  count(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned",
      "missing")
      )
  ) %>%
  tidyr::complete(concerned, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup() %>% 
  # Hacky need to fix later
  mutate(question = rep("", 8))

concerned_missing_data <-
  concerned_data %>% 
  filter(concerned == "missing") %>% 
  summarise(text = paste0(concerned, ": ", n))

concerned_data %>% 
  likert_plot_odd(.,
                  title = "Compared to what you think of as a typical researcher\nwho uses preexisting observational data in your field,\nhow concerned are you with research\n trustworthiness, bias, rigour, and reproducibility.",
                  questions = question,
                  levels = concerned,
                  n = n,
                  levels_to_drop = "missing",
                  positive_side = "lower",
                  text_push = 0.08,
                  limits = c(-0.5, 1),
                  legend_row = 3) +
  geom_text(concerned_missing_data, mapping = aes(x = 1, y = 1.1, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 5, l = 50),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# Number of studies published

The median of the number of studies published is `r median(raw$n_studies, na.rm = T)` with an IQR of `r IQR(raw$n_studies, na.rm = T)`. The maximum number of studies published is `r max(raw$n_studies, na.rm = T)` while the minimum is `r min(raw$n_studies, na.rm = T)`. There were `r raw %>% filter(is.na(n_studies)) %>% nrow(.)` missing responses to this question.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
raw %>% 
  select(n_studies) %>% 
  filter(!is.na(n_studies)) %>%
  ggplot() +
  aes(
    x = n_studies
  ) +
  geom_histogram() +
  labs(
    x = "The number of studies published",
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line()
  )
```

# Software used to analyse the data

There were `r raw %>% filter(is.na(programming_language)) %>% nrow(.)` missing responses to this question. The frequency of responses can be found below. Keep in mind that one respondent could choose multiple software languages.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
raw %>% 
  select(programming_language) %>% 
  filter(!is.na(programming_language)) %>% 
  separate_rows(programming_language, sep = ",") %>% 
  count(programming_language) %>% 
  rename(
    `Programming language` = programming_language,
    N = n) %>% 
  arrange(desc(N)) %>% 
  kableExtra::kbl()
```