### 2023-03-14 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# A survey of registration practices among observational researchers using preexisting datasets

<!-- badges: start -->
<!-- badges: end -->

The repository contains the data and the analysis for the research paper titled "A survey of registration practices among observational researchers using preexisting datasets". Other materials belonging to the project can be found on the project's OSF page at [link](link).

## Folder structure