### 2023-03-14 manuscript/interactive_report.Rmd

---
title: "Results exploration"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
# r_refs("references.bib")

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

what is this document...

# Subset

Subset data for all the results.

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
               min = min(processed$n_studies, na.rm = T), max = max(processed$n_studies, na.rm = T), value = max(processed$n_studies, na.rm = T), step = 1),
  radioButtons("methods", label = "Used the following methods:", choices = c("Blind", "Script", "Confirmatory", "Exploratory"), selected = NULL, inline = FALSE),
    radioButtons("concerned", label = "Concerned with rigour and reproducibility:", choices = c("Negative", "Neutral", "Support"), selected = NULL, inline = FALSE)
)

# Filter data
filtered_data <- reactive({
  processed %>% 
    filter(
      n_studies == input$n_studies
    )
  })
```

# Reproducibility and trustworthiness

The reproducible and trustworthiness ratings with the current ALSPAC system.

```{r}
# Prepare plot data
typically_plot_data <- 
  processed %>%
  select(contains("typically_")) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>% 
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

# Create likert package data not including the missing values
typically_plot_data <- likert(typically_plot_data)

# Create figure
typically_plot <- 
  plot(typically_plot_data)
```