### 2023-03-17 manuscript/interactive_report.Rmd

---
title: "Results exploration"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
library(plotly)
# r_refs("references.bib")

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

what is this document...

# Subset

Subset data for all the results.

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
              min = min(processed$n_studies, na.rm = T),
              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
  selectInput("methods", label = "Used the following methods:",
              choices = c(
                "Blind" = "method_blind",
                "Script" = "method_script",
                "Preregistration" = "method_preregistered",
                "Confirmatory" = "method_confirmatory", 
                "Exploratory" = "method_exploratory"),
              selected = NULL, multiple = TRUE),
  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
              choices = c(
                "Less concerned",
                "Neutral",
                "More concerned"),
              selected = c(
                "Less concerned",
                "Neutral",
                "More concerned"), multiple = TRUE),
  selectInput("language", label = "Programming language used for the analysis:",
              choices = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ),
              selected = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ), multiple = TRUE)
  )

# Filter data
filtered_data <- reactive({
  concerned_options <-list(
    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
    "Neutral" =  c("as concerned as a typical researcher in my field"),
    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
  )

  concerned_filter <- unlist(concerned_options[input$concerned], use.names = FALSE)

  processed %>% 
    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
  })
# select(filtered_data(), concerned)
# shiny::renderTable(filtered_data())
shiny::renderText({nrow(filtered_data())})
```

```{r echo=FALSE, warning=FALSE}
# Create typically trustworthy and reproducible plot
# Prepare plot data
typically_plot_data <- reactive({
  filtered_data() %>%
    select(contains("typically_")) %>%
    mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>%
    rename(
      Reproducible = typically_reproducible,
      Trustworthy = typically_trustworthy
      ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
typically_plot <- reactive({
  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right") +
  labs(title = "Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...")
})

# Create ecaw trustworthy and reproducible plot
# Prepare plot data
ecaw_plot_data <- reactive({
  filtered_data() %>%
  select(contains("ecaw_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more"))
    )) %>%
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
ecaw_plot <- reactive({
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right") +
  labs(title = "Compared to a typical study using preexisting observational data, a study using an ECAW would be...")
})

# Join the two plots together to make one figure
renderPlot({ typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2) })
```

```{r}
renderPlot({
# Create histogram of the number of observational studies done
filtered_data() %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram(binwidth = 3) +
  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  labs(
    x = "Number of studies",
    y = "Count",
    title = "Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?"
  ) +
  papaja::theme_apa()
})
```