### 2023-03-18 manuscript/interactive_report.Rmd
made interactive report prettier, added more setting options

---
title: "Results exploration"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
library(plotly)
library(DT)
# r_refs("references.bib")

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

The aim of this document is to allow users to interactively explore the descriptive results of the study titled _"A survey of registration practices among observational researchers using preexisting datasets"_.

# Settings

By using these filters you can subset the dataset for all the results presented in this report. The full dataset only contains the responses of those participants who finished the survey (n = `r nrow(processed)`).

```{r, echo=FALSE}
# Create a panel for the settings
inputPanel(
  # Filter by the number of studies
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
              min = min(processed$n_studies, na.rm = T),
              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
  # Filter by how concerned the respondent is with rigour and reproducibility
  # TODO: The last level was recorded with a capital letter, needs to be changed because it is a major source of error
  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
              choices = c("very much less concerned", "less concerned", "somewhat less concerned", "as concerned as a typical researcher in my field", "somewhat more concerned", "more concerned", "Very much more concerned", "missing"),
              selected = c("very much less concerned", "less concerned", "somewhat less concerned", "as concerned as a typical researcher in my field", "somewhat more concerned", "more concerned", "Very much more concerned", "missing"),
              multiple = TRUE),
  # Filter by programming language used for the analysis
  selectInput("language", label = "Programming language used for the analysis:",
              choices = c("R", "Stata", "SPSS", "SAS", "Python", "Mplus", "Bash", "MATLAB", "Nextflow", "plink2", "irrelevant"),
              selected = c("R", "Stata", "SPSS", "SAS", "Python", "Mplus", "Bash", "MATLAB", "Nextflow", "plink2", "irrelevant"),
              multiple = TRUE),
  br(),
  # Filtered by the methods used by the respondents
  ## BLinded analysis
  selectInput("blinded", label = "Blinded data analyst",
              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              multiple = TRUE),
  ## Preregistration
  selectInput("prereg", label = "Preregistration",
              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              multiple = TRUE),
  ## Script
  selectInput("script", label = "Analysis script",
              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              multiple = TRUE),
  ## Confirmatory analysis
  selectInput("confirmatory", label = "Confirmatory analysis",
              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              multiple = TRUE),
  ## Exploratory analysis
  selectInput("exploratory", label = "Exploratory analysis",
              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
              multiple = TRUE)
  )

# Filter data based on the settings
filtered_data <- reactive({
  # Create a filter for concerned and methods that only includes response levels
  # TODO: Not the most elegant solution...
  concerned_filter <- input$concerned[input$concerned != "missing"]
  blinded_filter <- input$blinded[input$blinded != "Missing"]
  prereg_filter <- input$prereg[input$prereg != "Missing"]
  script_filter <- input$script[input$script != "Missing"]
  confirmatory_filter <- input$confirmatory[input$confirmatory != "Missing"]
  exploratory_filter <- input$exploratory[input$exploratory != "Missing"]

  res <-
    processed %>% 
    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
    # Filter concerned by response levels and keep all missing responses for now
    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
    # Filter methods by levels and keep all missing responses for now
    filter(method_blind %in% blinded_filter | is.na(method_blind)) %>% 
    filter(method_preregistered %in% prereg_filter | is.na(method_preregistered)) %>% 
    filter(method_script %in% script_filter | is.na(method_script)) %>% 
    filter(method_confirmatory %in% confirmatory_filter | is.na(method_confirmatory)) %>% 
    filter(method_exploratory %in% exploratory_filter | is.na(method_exploratory)) %>% 
    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))

  # Filter missing responses from concerned and method if specified
  # TODO: Not the most elegant solution...
  if ("missing" %ni% input$concerned) res <- dplyr::filter(res, !is.na(concerned))
  if ("Missing" %ni% input$blinded) res <- dplyr::filter(res, !is.na(method_blind))
  if ("Missing" %ni% input$prereg) res <- dplyr::filter(res, !is.na(method_preregistered))
  if ("Missing" %ni% input$script) res <- dplyr::filter(res, !is.na(method_script))
  if ("Missing" %ni% input$confirmatory) res <- dplyr::filter(res, !is.na(method_confirmatory))
  if ("Missing" %ni% input$exploratory) res <- dplyr::filter(res, !is.na(method_exploratory))

  # Return filtered data
  return(res)
  })
```

# Dataset

```{r echo=FALSE}
# Get the index of open ended question for condensing the long text
condensed_cols <- which(colnames(processed) %in% c("unsure_explain", "drawback_ecaw", "suggestions", "comments", "contact")) - 1

# Render the dataset as a pretty table
DT::renderDataTable(filtered_data(), rownames = FALSE, options = list(
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20),
  scrollX = TRUE,
  autoWidth = TRUE,
  columnDefs = list(
    list(
      className = "dt-center", targets = 0
    ),
    list(
      targets = condensed_cols,
      render = htmlwidgets::JS(
        "function(data, type, row, meta) {",
        "return type === 'display' && data != null && data.length > 25 ?",
        "'<span title=\"' + data + '\">' + data.substr(0, 25) + '...</span>' : data;",
        "}"
      )
    )
  )
))
```

# Results from the manuscript

```{r echo=FALSE, warning=FALSE}
# Create typically trustworthy and reproducible plot
# Prepare plot data
typically_plot_data <- reactive({
  filtered_data() %>%
    select(contains("typically_")) %>%
    mutate(
      across(
        .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree"
        ))
      )
    ) %>%
    rename(
      Reproducible = typically_reproducible,
      Trustworthy = typically_trustworthy
    ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
typically_plot <- reactive({
  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.5) +
    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
    labs(title = str_wrap("Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...", width = 50)) +
    theme(
      title = element_text(size = 14),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      legend.text = element_text(size = 13.5),
      legend.title = element_blank()
    )
})

# Create ecaw trustworthy and reproducible plot
# Prepare plot data
ecaw_plot_data <- reactive({
  filtered_data() %>%
    select(contains("ecaw_")) %>%
    # Transform not wanted response values to NA
    # likert::likert drops NA values silently
    # When var transformed to factor these values would be transformed to NA
    # Automatically but I try to be explicit
    mutate(
      across(
        everything(),
        ~ case_when(
          . == "I don't understand the question" ~ NA_character_,
          TRUE ~ .
        )
      )
    ) %>%
    mutate(
      across(
        .fns = ~ factor(., levels = c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more"
        ))
      )
    ) %>%
    rename(
      Reproducible = ecaw_reproducible,
      Trustworthy = ecaw_trustworthy
    ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
ecaw_plot <- reactive({
  plot(ecaw_plot_data(), digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.5) +
    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
    labs(title = str_wrap("Compared to a typical study using preexisting observational data, a study using an ECAW would be...", width = 50)) +
    theme(
      title = element_text(size = 14),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      legend.text = element_text(size = 13.5),
      legend.title = element_blank(),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
    )
})

# Join the two plots together to make one figure
renderPlot({
  typically_plot() + ecaw_plot() + plot_layout(ncol = 1, nrow = 2)
})
```

---

```{r echo=FALSE}
# Used methods figures
method_plot_data <- reactive({
  filtered_data() %>%
    select(starts_with("method_")) %>%
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>%
    group_by(type) %>%
    count(scale) %>%
    ungroup() %>%
    replace_na(list(scale = "Missing")) %>%
    mutate(
      scale = factor(scale, levels = c(
        "Never or almost never",
        "Sometimes",
        "About half the time",
        "Most of the time",
        "Always or almost always",
        "Missing",
        "I don't understand the question"
      )),
      type = case_when(
        type == "blind" ~ "Blind the data analyst",
        type == "confirmatory" ~ "Contain confirmatory analysis",
        type == "exploratory" ~ "Contain exploratory analysis",
        type == "preregistered" ~ "Are preregistered",
        type == "script" ~ "Share analysis scripts",
      ),
      type = factor(type, levels = c(
        "Contain exploratory analysis",
        "Contain confirmatory analysis",
        "Blind the data analyst",
        "Share analysis scripts",
        "Are preregistered"
      ))
    ) %>%
    group_by(type) %>%
    tidyr::complete(scale, fill = list(n = 0)) %>%
    mutate(
      percentage = n / sum(n)
    ) %>%
    ungroup()
})

renderPlot({
  method_plot_data() %>%
    # filter(scale %ni% c("Missing", "I don't understand the question")) %>%
    ggplot() +
    aes(
      x = percentage,
      y = type,
      fill = scale
    ) +
    geom_bar(position = "stack", stat = "identity") +
    scale_x_continuous(
      limits = c(0, 1),
      labels = scales::label_percent()
    ) +
    scale_fill_viridis(option = "A", discrete = TRUE) +
    scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
    labs(
      x = "Percentage",
      fill = "Response",
      title = "The studies using preexisting observational data that I am involved in..."
    ) +
    papaja::theme_apa() +
    theme(
      title = element_text(size = 17),
      axis.text = element_text(size = 15),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.text = element_text(size = 14),
      legend.title = element_blank(),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
    )
})
```

---

```{r echo=FALSE}
# Prepare plot data
alspac_plot_data <- reactive({
  filtered_data() %>%
    select(contains("alspac_")) %>%
    # Transform not wanted response values to NA
    # likert::likert drops NA values silently
    # When var transformed to factor these values would be transformed to NA
    # Automatically but I try to be explicit
    mutate(
      across(
        everything(),
        ~ case_when(
          . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
          TRUE ~ .
        )
      )
    ) %>%
    mutate(
      across(
        .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree"
        ))
      )
    ) %>%
    rename(
      `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
      `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
      `ALSPAC should run a study on ECAWs.` = alspac_study,
      `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
    )
})

# Create likert package data not including the missing values
alspac_plot_likert_data <- reactive({
  likert(alspac_plot_data())
})

# Create figure
renderPlot({
  plot(alspac_plot_likert_data(), digits = 1, ordered = TRUE, group.order = names(alspac_plot_data()), text.size = 5.5) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 25), limits = rev(names(alspac_plot_data))) +
    labs(title = str_wrap("Thinking about a study you may run with ALSPAC data (or one that you have recently run)..."), width = 25) +
    theme(
      axis.title.x = element_blank(),
      title = element_text(size = 14),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 16),
      legend.text = element_text(size = 15),
      legend.title = element_blank(),
      plot.margin = margin(t = 0, r = 0, b = 3.5, l = 0, unit = "lines"),
      legend.position = c(.35,-.2),
      legend.direction = "horizontal"
    )
})
```

# Results from supplementary materials C

```{r echo=FALSE}
renderPlot({
  # Create histogram of the number of observational studies done
  filtered_data() %>%
    # Exclude missing responses
    filter(!is.na(n_studies)) %>%
    ggplot() +
    aes(x = n_studies) +
    geom_histogram(binwidth = 3) +
    # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
    labs(
      x = "Number of studies",
      y = "Count",
      title = str_wrap("Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?", width = 70)
    ) +
    papaja::theme_apa() +
    theme(
      title = element_text(size = 15),
      axis.text = element_text(size = 15),
      axis.title = element_text(size = 17)
    )
})
```

---

```{r echo=FALSE}
# Programming languages table
renderTable({
  denominator <- nrow(filtered_data())

  filtered_data() %>%
    rename(lang = programming_language) %>%
    separate_rows(lang, sep = ",") %>%
    count(lang) %>%
    replace_na(list(lang = "Missing")) %>%
    # Removing irrelevant
    filter(lang != "irrelevant") %>%
    ungroup() %>%
    mutate(
      `Percentage of respondents` = round(n / denominator * 100, 2)
    ) %>%
    arrange(desc(n)) %>%
    rename(
      `Programming language` = lang,
      `N` = n
    )
})
```

---

```{r echo=FALSE}
# Concerned figure data preparation
concerned_plot_data <- reactive({
  filtered_data() %>%
    select(concerned) %>%
    mutate(
      concerned = tolower(concerned),
      concerned = factor(concerned, levels = c(
        "very much less concerned",
        "less concerned",
        "somewhat less concerned",
        "as concerned as a typical researcher in my field",
        "somewhat more concerned",
        "more concerned",
        "very much more concerned"
      ))
    ) %>%
    filter(!is.na(concerned))
})

# Create likert package data not including the missing values
concerned_likert_plot_data <- reactive({
  likert(concerned_plot_data())
})

# Create figure
renderPlot({
  plot(concerned_likert_plot_data(), digits = 1, text.size = 6, group.order = names(concerned_plot_data())) +
    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
    labs(title = str_wrap("Compared to what you think of as a typical researcher who uses preexisting observational data in your field, how concerned are you with research trustworthiness, bias, rigour, and reproducibility ...", width = 100)) +
    theme(
      title = element_text(size = 14),
      axis.text.y = element_blank(),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      legend.text = element_text(size = 11),
      legend.title = element_blank()
    )
})
```