### 2023-04-28 manuscript/supplementary_materials.Rmd

# Supplementary Material A. Deviations from the preregistration.

\begin{enumerate}

  \item We rephrased our second objective to be more accurate. The preregistration reads “2.	To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices (survey Blocks 3-6). This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability. ”. The manuscript reads “(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.”
  \item Marton Kovacs was added as a contributor during data collection. This led to a few of the projected contributor roles outlined in the preregistration to be different from the final contributor roles outlined in the manuscript.
  \item The preregistration stated that: “We will tabulate descriptive summary statistics for all the survey questions.” Instead of tabulating the results, the manuscript presents this data in figures. We feel that the figures are easier to digest as compared to tabulated data.
  \item The preregistration stated that: “we   will   present   results   that   include   responses   from participants who did not complete the entire survey, alongside the associated response rate for each question.”. The results from all participants are presented in Supplementary Material D. However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed)` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(processed_all) / 1094 * 100)`% for at least partially complete surveys.”
  \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
  \item We performed a brief visual inspection of the data presented in spreadsheet format. Three participants clicked through to the end of the survey—so they were coded as completing the survey—however, they only responded to a few of the first questions. One other participant appears to have provided non-sincere responses in that all their responses were neutral or skipped and they completed the survey in 146 seconds. We retained these four participants as partially completed responses (results in Supplementary Material D), but removed them from the main dataset. These decisions were not preregistered.
  \item We did not preregister which results we would present in the abstract. We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only `r pull(filter(study, support == "negative"), percentage)`\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as `r pull(filter(willing, support == "positive"), percentage)`\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”

\end{enumerate}

\pagebreak

# Supplementary Material B. Invitation email
## Original email

Dear ALSPAC Data User,

We are working with Dr Robert Thibault, a postdoctoral scholar at Stanford University and the University of Bristol, to support his work on scientific rigour and reproducibility in observational studies.

To assist him in his research we would be very grateful if you would consider completing a short survey. The purpose of this is to understand researcher’s practices and thoughts regarding the rigour and reproducibility of observational research that uses pre-existing datasets (such as the ALSPAC resource). Results from this survey may be used in the future to inform initiatives for ALSPAC to better serve our users and to maximise the quality of the research using ALSPAC.

The survey has 21 questions across six sections. It will be open until November 1 at this link: https://bristolexppsych.eu.qualtrics.com/jfe/form/SV_3mVgk4lzXx4kp02. It will take approximately 10-20 minutes to complete. The data you provide is completely anonymous and further information is available on the consent form at the start of the survey.

If you have any questions or comments, please send them directly to Robert (robert.thibault@bristol.ac.uk).

Kind regards,

The ALSPAC Executive

---

## First follow-up email (sent after 1 week)

Dear ALSPAC data user,

Thank you very much to those of you who have already completed the survey below.
There is still time to complete the survey (see details in the original email below), which closes on November 1st.

Please send any comments or queries directly to Robert (robert.thibault@bristol.ac.uk).

Many thanks,

The ALSPAC Executive

---

## Second follow-up email (sent after 2 weeks)

Dear ALSPAC data user,

Thank you very much to those of you who have already completed the survey below.
This will be our final email inviting you to complete the survey (see details in the original email below), which closes of November 1st.

The median time to complete the survey has been 8 minutes.

Please send any comments or queries directly to Robert (robert.thibault@bristol.ac.uk).

Many thanks,
The ALSPAC Executive

\pagebreak

# Supplementary Material C. Participant Characteristics

```{r languageTable}
# Programming languages table
# We use the number of all participants who responded to the questionnaire as the denominator
denominator <- nrow(processed)

processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  replace_na(list(lang = "Missing")) %>% 
  # Removing irrelevant
  filter(lang != "irrelevant") %>% 
  ungroup() %>% 
  mutate(
    `Percentage of respondents` = round(n / denominator * 100, 0)
  ) %>% 
  arrange(desc(n)) %>% 
  rename(
    `Programming language` = lang,
    `N` = n
  ) %>% 
  kbl(
  .,
  format = "latex",
  col.names = c("Programming language", "N", "Percentage of respondents"),
  booktabs = T,
  escape = F,
  caption = "(ref:table1c-caption)",
  centering = T,
  position = "H") %>%
  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
  row_spec(row = 0, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:table1c-caption) __Table C1. “What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)”__
```{=latex}
\end{singlespace}
}
```

```{r numStudiesPlot, fig.align="center", fig.cap="(ref:numStudiesPlotCaption)", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
# Create text for missing responses
n_studies_missing_text <- 
  processed %>% 
  filter(is.na(n_studies)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create histogram of the number of observational studies done
processed %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram(binwidth = 3) +
  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  labs(
    x = "Number of publications",
    y = "Number of respondents",
    title = "Approximately how many studies have you published using\na preexisting observational dataset (e.g., the ALSPAC dataset)?"
  ) +
  papaja::theme_apa()

# ggplot() +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies == 0),
#     mapping = aes(x = n_studies), binwidth = 1, color = 'red') +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies != 0),
#     mapping = aes(x = n_studies), binwidth = 5) +
#   # annotate(geom = "text", label = n_studies_missing_text, y = 30) +
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
#   labs(
#     x = "Number of observational studies done prior using ALSPAC",
#     y = "Count"
#   ) +
#   papaja::theme_apa()
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:numStudiesPlotCaption) __Figure C1. Responses to survey question about number of relevant publications.__
```{=latex}
\end{singlespace}
}
```

```{r concernedPlot, warning=FALSE, fig.cap="(ref:concernedPlotCaption)", fig.align="center", out.width="100%", fig.width=10, fig.height=4, fig.path='figs/', dev=c('png', 'pdf')}
# Concerned figure data preparation
concerned_plot_data <-
  processed %>% 
  select(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  )

# Check distinct values in data
# distinct(concerned_plot_data, concerned)

# Create missing values text
concerned_missing_n <- 
  concerned_plot_data %>% 
  filter(is.na(concerned)) %>% 
  summarise(
    n = n()
  ) %>% 
  pull(n)

concerned_plot_data <- filter(concerned_plot_data, !is.na(concerned))

# Create likert package data not including the missing values
concerned_likert_plot_data <- likert(concerned_plot_data)

# Create figure
plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = str_wrap("Compared to what you think of as a typical researcher who uses preexisting observational data in your field, how concerned are you with research trustworthiness, bias, rigour, and reproducibility ...", width = 100)) +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 10),
    legend.title = element_blank()
  )
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:concernedPlotCaption) __Figure C2. Responses to survey question about number of relevant publications.__ We excluded missing responses from this figure (_n_ = `r concerned_missing_n`).
```{=latex}
\end{singlespace}
}
```

\pagebreak

# Supplementary Material D. Figures and Tables including participants who partially complete the survey
## Participants

```{r languageDescriptivesAll, include=FALSE}
# List of used programming languages and their counts text
language_text_all <- 
  processed_all %>%
  rename(lang = programming_language) %>%
  separate_rows(lang, sep = ",") %>%
  count(lang) %>%
  filter(!is.na(lang)) %>%
  arrange(desc(n)) %>%
  # Removing irrelevant
  filter(lang != "irrelevant") %>%
  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>%
  pull(lang_n) %>%
  glue::glue_collapse(., sep = ", ", last = ", and ")
```

```{r concernedDescriptivesAll, include=FALSE}
# Calculate concerned percentage
concerned_all <- support_percentage(
  processed_all,
  concerned,
  c(
    "very much less concerned",
    "less concerned",
    "somewhat less concerned",
    "as concerned as a typical researcher in my field",
    "somewhat more concerned",
    "more concerned",
    # This level label starts with a capital letter
    "Very much more concerned"
  ),
  exclude_missing = TRUE
)
```

Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using pre-existing observational data. They reported using the following programming languages or software packages: `r language_text_all` \footnote[1]{Participants could select multiple responses to this survey question.}. `r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses pre-existing observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned.

## Survey results

```{r typicallyEcawDescriptivesAll, include=FALSE}
# Calculate typical trustworthiness and reproducibility ratings
typically_trustworthy_all <- support_percentage(
  processed_all,
  typically_trustworthy,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  "I don't understand the question"
)

typically_reproducible_all <- support_percentage(
  processed_all,
  typically_reproducible,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  "I don't understand the question"
)

# Calculate ecaw trustworthiness and reproducibility ratings
ecaw_trustworthy_all <- support_percentage(
  processed_all,
  ecaw_trustworthy,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"
  ),
  "I don't understand the question"
)

ecaw_reproducible_all <- support_percentage(
  processed_all,
  ecaw_reproducible,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"
  ),
  "I don't understand the question"
)
```

Most respondents agreed that studies that analyze pre-existing observational datasets are trustworthy (`r pull(filter(typically_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy_all, support == "positive"), n)`/`r pull(filter(typically_trustworthy_all, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible_all, support == "positive"), n)`/`r pull(filter(typically_reproducible_all, support == "positive"), n_sum)`) (Supplementary Figure D1, top panel). At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy_all, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using pre-existing observational data (Supplementary Figure D1, bottom panel).

```{r typicallyEcawPlotAll, warning=FALSE, fig.cap="(ref:typicallyEcawPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed_all %>%
#   select(contains("typically_")) %>%
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question")

# Calculate missing responses
typically_missing_all <-
  processed_all %>%
  select(contains("typically_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  )

# Prepare plot data
typically_plot_data_all <-
  processed_all %>%
  select(contains("typically_")) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

# Create likert package data not including the missing values
typically_plot_data_all <- likert(typically_plot_data_all)

# Create figure
typically_plot_all <-
  plot(typically_plot_data_all, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank()
  )

# Create ecaw trustworthy and reproducible plot
# Calculate missing responses
ecaw_missing_all <-
  processed_all %>%
  select(contains("ecaw_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  )

# Prepare plot data
ecaw_plot_data_all <-
  processed_all %>%
  select(contains("ecaw_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Much less",
        "Somewhat less",
        "About the same",
        "Somewhat more",
        "Much more"
      ))
    )
  ) %>%
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  )

# Create likert package data not including the missing values
ecaw_plot_data_all <- likert(ecaw_plot_data_all)

# Create figure
ecaw_plot_all <-
  plot(ecaw_plot_data_all, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
  )

# Join the two plots together to make one figure
typically_plot_all + ecaw_plot_all + plot_layout(ncol = 1, nrow = 2)
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:typicallyEcawPlotAllCaption) __Figure D1. Responses to the survey questions on trustworthiness and reproducibility of observational research with pre-existing data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust”. The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel excluded missing responses (n = 0; `r pull(filter(typically_missing_all, item == "Reproducible" & value == "Missing"), n)`) and had no selection of the option _“I don’t understand the question”_. The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "I don't understand the question"), n)`).
```{=latex}
\end{singlespace}
}
```

```{r methodDescriptivesAll, include=FALSE}
# Define the response level order for these variables
method_levels <- c(
  "Never or almost never",
  "Sometimes",
  "About half the time",
  "Most of the time",
  "Always or almost always",
  "I don't understand the question",
  "Missing"
)

# Calculate the frequency and proportion of responses for each response level
confirmatory_all <- level_percentage(
  processed_all,
  method_confirmatory,
  method_levels,
  exclude_missing = FALSE
) %>%
  filter(method_confirmatory %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
  summarise(
    n = sum(n),
    # Since n_sum should be the same for the whole dataset we can use unique to return one value
    n_sum = unique(n_sum),
    percentage = sum(percentage))

exploratory_all <- level_percentage(
  processed_all,
  method_exploratory,
  method_levels,
  exclude_missing = FALSE
) %>%
  dplyr::filter(method_exploratory  %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
  summarise(
    n = sum(n),
    # Since n_sum should be the same for the whole dataset we can use unique to return one value
    n_sum = unique(n_sum),
    percentage = sum(percentage))

preregistered_all <- level_percentage(
  processed_all,
  method_preregistered,
  method_levels,
  exclude_missing = FALSE
)

blind_all <- level_percentage(
  processed_all,
  method_blind,
  method_levels,
  exclude_missing = FALSE
)

script_all <- level_percentage(
  processed_all,
  method_script,
  method_levels,
  exclude_missing = FALSE
)
```

Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Supplementary Figure 2D). About half reported sharing their analysis scripts never or almost never (`r filter(script_all, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script_all, method_script == "Never or almost never") %>% pull(n)`/`r filter(script_all, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script_all, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script_all, method_script == "Sometimes") %>% pull(n)`/`r filter(script_all, method_script == "Sometimes") %>% pull(n_sum)`). `r filter(blind_all, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both confirmatory (`r pull(confirmatory_all, percentage)`%; `r pull(confirmatory_all, n)`/`r pull(confirmatory_all, n_sum)`) and exploratory (`r pull(exploratory_all, percentage)`%; `r pull(exploratory_all, n)`/`r pull(exploratory_all, n_sum)`) analyses at least sometimes.

```{r methodPlotAll, warning=FALSE, fig.cap="(ref:methodPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
# Used methods figures
method_plot_data_all <-
  processed_all %>%
  select(starts_with("method_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "scale",
  ) %>%
  separate(variable, into = c("name_prefix", "type")) %>%
  group_by(type) %>%
  count(scale) %>%
  ungroup() %>%
  replace_na(list(scale = "Missing")) %>%
  mutate(
    scale = factor(scale, levels = c(
      "Never or almost never",
      "Sometimes",
      "About half the time",
      "Most of the time",
      "Always or almost always",
      "Missing",
      "I don't understand the question"
    )),
    type = case_when(
      type == "blind" ~ "Blind the data analyst",
      type == "confirmatory" ~ "Contain confirmatory analysis",
      type == "exploratory" ~ "Contain exploratory analysis",
      type == "preregistered" ~ "Are preregistered",
      type == "script" ~ "Share analysis scripts",
    ),
    type = factor(type, levels = c(
      "Contain exploratory analysis",
      "Contain confirmatory analysis",
      "Blind the data analyst",
      "Share analysis scripts",
      "Are preregistered"
    ))
  ) %>%
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>%
  mutate(
    percentage = n / sum(n)
  ) %>%
  ungroup()

method_plot_data_all %>%
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>%
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
  labs(
    title = "The studies using preexisting observational data that I am involved in..."
  ) +
  papaja::theme_apa() +
  theme(
    title = element_text(size = 17),
    axis.text = element_text(size = 15),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.text = element_text(size = 14),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
  )
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:methodPlotAllCaption) __Figure D2. Responses to survey questions about the research practices of participants.__
```{=latex}
\end{singlespace}
}
```

```{r aslpacDescriptivesAll, include=FALSE, warning=FALSE, message=FALSE}
willing_all <- support_percentage(
  processed_all,
  alspac_less_willing,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

opt_all <- support_percentage(
  processed_all,
  alspac_opt_in,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

study_all <- support_percentage(
  processed_all,
  alspac_study,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)

prefer_all <- support_percentage(
  processed_all,
  alspac_prefer_ecaw,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"
  ),
  c(
    "I don't understand the question",
    "Unsure"
  )
)
```

`r pull(filter(willing_all, support == "positive"), percentage)`% (`r pull(filter(willing_all, support == "positive"), n)`/`r pull(filter(willing_all, support == "positive"), n_sum)`) of respondents agreed (versus `r pull(filter(willing_all, support == "negative"), percentage)`%; `r pull(filter(willing_all, support == "negative"), n)`/`r pull(filter(willing_all, support == "negative"), n_sum)` who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Supplementary Figure 3D). `r pull(filter(opt_all, support == "positive"), percentage)`% (`r pull(filter(opt_all, support == "positive"), n)`/`r pull(filter(opt_all, support == "positive"), n_sum)`) agreed (`r pull(filter(opt_all, support == "negative"), percentage)`%; `r pull(filter(opt_all, support == "negative"), n)`/`r pull(filter(opt_all, support == "negative"), n_sum)` disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. `r pull(filter(study_all, support == "positive"), percentage)`% (`r pull(filter(study_all, support == "positive"), n)`/`r pull(filter(study_all, support == "positive"), n_sum)`) agreed (`r pull(filter(study_all, support == "negative"), percentage)`%; `r pull(filter(study_all, support == "negative"), n)`/`r pull(filter(study_all, support == "negative"), n_sum)` disagreed) that ALSPAC should run a study on ECAWs. `r pull(filter(prefer_all, support == "positive"), percentage)`% (`r pull(filter(prefer_all, support == "positive"), n)`/`r pull(filter(prefer_all, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer_all, support == "negative"), n)`/`r pull(filter(prefer_all, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using conventional preregistration.

```{r alspacPlotAll, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
# Check if there is any missing and don't understand responses separately
alspac_missing_all <-
  processed_all %>%
  select(contains("alspac_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
  mutate(
    item = str_to_title(str_remove(item, "^[^_]*_")),
    # We reorder it by making it a factor
    item = factor(item, levels = c(
      "Less_willing",
      "Opt_in",
      "Study",
      "Prefer_ecaw"
    )),
    # By making it a factor we can complete missing levels in the data later
    value = factor(value, levels = c(
      "Missing",
      "I don't understand the question",
      "Unsure"
    ))
  ) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  ) %>% 
  # Add n = 0 if no one choose a specific factor level
  tidyr::complete(value, fill = list(n = 0))

# Prepare plot data
alspac_plot_data_all <-
  processed_all %>%
  select(contains("alspac_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
    `ALSPAC should run a study on ECAWs.` = alspac_study,
    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
  )

# Create likert package data not including the missing values
alspac_plot_likert_data_all <- likert(alspac_plot_data_all)

# Create figure
plot(alspac_plot_likert_data_all, digits = 1, text.size = 6) +
  # TODO: have to check if this is not messing with the results
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30), limits = rev(names(alspac_plot_data))) +
  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
  theme(
    axis.title.x = element_blank(),
    title = element_text(size = 15.5),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.title = element_blank()
  )
```

```{=latex}
{\smaller[1]
\begin{singlespace}
```
(ref:alspacPlotAllCaption) __Figure D3. Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Unsure"), n)`). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study”.
```{=latex}
\end{singlespace}
}
```

\pagebreak

# Supplementary Material E. Description of ALSPAC study numbers

Pregnant women resident in Avon, UK with expected dates of delivery between 1st April 1991 and 31st December 1992 were invited to take part in the study.  20,248 pregnancies have been identified as being eligible and the initial number of pregnancies enrolled was 14,541. Of the initial pregnancies, there was a total of 14,676 foetuses, resulting in 14,062 live births and 13,988 children who were alive at 1 year of age. When the oldest children were approximately 7 years of age, an attempt was made to bolster the initial sample with eligible cases who had failed to join the study originally. As a result, when considering variables collected from the age of seven onwards (and potentially abstracted from obstetric notes) there are data available for more than the 14,541 pregnancies mentioned above: The number of new pregnancies not in the initial sample (known as Phase I enrolment) that are currently represented in the released data and reflecting enrolment status at the age of 24 is 906, resulting in an additional 913 children being enrolled (456, 262 and 195 recruited during Phases II, III and IV respectively). The phases of enrolment are described in more detail in the cohort profile paper and its update (see footnote 5 below). The total sample size for analyses using any data collected after the age of seven is therefore 15,447 pregnancies, resulting in 15,658 foetuses. Of these 14,901 children were alive at 1 year of age.

Of the original 14,541 initial pregnancies, 338 were from a woman who had already enrolled with a previous pregnancy, meaning 14,203 unique mothers were initially enrolled in the study. As a result of the additional phases of recruitment, a further 630 women who did not enrol originally have provided data since their child was 7 years of age. This provides a total of 14,833 unique women (G0 mothers) enrolled in ALSPAC as of September 2021.

G0 partners were invited to complete questionnaires by the mothers at the start of the study and they were not formally enrolled at that time. 12,113 G0 partners have been in contact with the study by providing data and/or formally enrolling when this started in 2010. 3,807 G0 partners are currently enrolled.